Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-04 trunk/doc/thesis/report.tex
chapter 1 reviewed

\section{Importance of Study}
\paragraph{}
The global electrical grid one the bases that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man. \cite{ref1}. It a widespread machine, distributed but interconnected system that provides power to households, companies and other organizations and therefore the world's economy is counting on it. The electrical grid knows numerous problems such as blackouts, current instabilities that costs billions of dollars yearly to both the utilities, utilities and insurances. The grid suffers from the problem of peak hours where energy consumption where in many case the one single utility cannot provide such energy.
\paragraph{}
The \gls{sg} comes to provide solutions to the problems that suffers from the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and is and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arises in the traditional grid such as the energy loss in the energy transmission lines and also in the overload of the electrical lines.
\paragraph{}
Another revolution is about to begin and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously without the guidance of humans to performs their tasks. Cisco predicts that the number of devices to be connected in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion. That make 6.58 devices connected for each person in the world whereas the number of devices connected in 2010 12.5 billion. This clearly shows that the shift to the \gls{iot} has already started.
\paragraph{}
\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}. The integration of the \gls{sg} and and \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The smart grid as a set of philosophical ideas counts on a set of novel functionalities. Once of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa. This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information. Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the Internet of things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible. This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}. From the consumer's side or more precisely the household's side, there are smart home energy management software in the market nowadays but none of these products offers integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication. Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
\paragraph{}
As \gls{sg} counts on numerous technologies to achieve its main goals. This usage of different technologies imposes another problem which is the heterogeneity of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. This means that there is a need to build a software within the whole system that would constitute an important part of it and will take care of making the communication and processes technology-transparent and give the possibility to all components of the system to work hand-in-hand in order their common goals independently of the data link layer technology (Ethernet, Wifi or Zigbee) they are counting on, the network layer technology (\gls{ipv4} or \gls{ipv6}).

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and also solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of things by making it communicate with any device in the Internet and having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational and energy rich.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility. To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}. This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}. This communication will go in the two directions and based on that the two-way communication will have to be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component. This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
The thesis aims to solves the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate one cannot solve without having to solve the other. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other network only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem. This is why the two-way communication was created and then the heterogeneity problem had to be solved in order to see everything working.
The system is divided into four components:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The \gls{tinyos} program for sensor transmission and appliance request handler
\item The heterogeneity handler and middleman gateway server
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four component exist to solve part of the problem. Each of these components will be discussed extensively in the Chapter 4 from the architectural aspect of the component and in Chapter 5 from the implementation level of the system. 
\paragraph{}
It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation point in Chapter 5. The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically. The consumption of data by is shown by external nodes in the \gls{iot} sending request to the sensor node in the \gls{wsn} in order to control and appliance's status i.e. turn on the television. This will prove the consumption of information by the nodes in the \gls{iot}.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world. This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.

@@

\chapter{Findings}
\paragraph{}
The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system. Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}. To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services. The purpose of this chapter is to measure the performance of the system. It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
\section{Experiment}
\paragraph{}
To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system. \gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it. This means that there is more priority given to routed packet that the processed ones. As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
\paragraph{Experiment 1:}
We will first have one mote in the \gls{wsn} that will be interacting with the outside world. We will start sending ping command periodically (once every 1 second) and record the delay and the jitter. Once every two minutes, we will add another mote to the system and continue pinging the same mote. We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system. This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
\paragraph{Experiment 2: }
The second experience is meant to measure the time it takes to transform the packet. In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system. The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
\section{Data Gathering}
\paragraph{}
After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
\paragraph{Experiment 1}
In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes. The experiment as a whole took 12 minutes and 720 pings we done. Table \ref{table:exp1} shows the result of the experiment.

\begin{table}[htbp]
    \begin{tabular}{lll}
    \hline
    number of motes & Delay (ms) & Jitter (ms) \\\hline
    1               & 87.5         & 1.2         \\ 
    2               & 88.1         & 2.0         \\
    3               & 88.3        & 2.7         \\
    4               & 88.8         & 3.1         \\
    5               & 90.0         & 3.3         \\
    6               & 89.5         & 3.5         \\
    \end{tabular}
    \caption{Experiment 1 delay and jitter}
    \label{table:exp1}
\end{table}
For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row. The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval. The result is displayed in the table \ref{table:exp1}.

\paragraph{Experiment 2}
The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system. The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient. Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.

@@

The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nano seconds. Depending on the machine's load.

@@

@@

 nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided by the 

@@

\backmatter
\singlespacing
\addcontentsline{toc}{section}{References}
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}

\doublespacing
\appendix
\chapter{How to install the system}
\section{Hardware needed}
In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services. The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi. A smartphone or a separate computer with Android emulator may de used.

>>>>

\section{Importance of Study}
\paragraph{}
The global electrical grid is one the bases that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man. \cite{ref1}. It a widespread machine, distributed but an interconnected system that provides power to households, companies and other organisations and therefore the world's economy is counting on it. The electrical grid knows numerous problems such as blackouts, current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
\paragraph{}
The \gls{sg} comes as a solution to the problems that suffers from the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from long energy transmission lines and also from the overload of the electrical lines.
\paragraph{}
Another revolution began and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously, without the guidance of humans, to performs their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion. That makes 6.58 devices connected for each person in the world whereas in 2010, the number of devices connected was 12.5 billion. This clearly shows that the shift to the \gls{iot} has already started.
\paragraph{}
\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid as a set of philosophical ideas counts on a set of novel functionalities. One of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa. This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information. Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which make the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deal with such heterogeneity and take the most out of it in order to make the communication as smooth and transparent as possible. This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's side or more precisely the household's side, there are smart home energy management systems in the market nowadays but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need to first build this two-way communication, to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely with this infrastructure. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This usage of different technologies brings another problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their common goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility. To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}. This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}. This communication will go in the two directions and based on that, the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component. This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
The thesis aims to solves the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate, one cannot solve without having to solve the other. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other networks only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem. This is why the two-way communication was created but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The \gls{tinyos} program for sensor transmission and appliance request handler
\item The heterogeneity handler and middleman gateway server
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four component is there to solve part of the problem. Each of these subsystems will be discussed extensively in the Chapter 4 from the architectural aspect of the component, and in Chapter 5 from the deployment point of view. 
\paragraph{}
It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically. The consumption of data by is shown by external nodes in the \gls{iot} sending requests to the sensor node in the \gls{wsn} in order to control the appliance's status i.e. turn On the television. This will prove the consumption of information by the nodes in the \gls{iot}.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world. This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.

@@

\chapter{Findings}
\paragraph{}
The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system. Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}. To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services. The purpose of this chapter is to measure the performance of the system. It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
\section{Experiment}
\paragraph{}
To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system. \gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it. This means that there is more priority given to routed packet that the processed ones. As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
\paragraph{Experiment 1:}
We will first have one mote in the \gls{wsn} that will be interacting with the outside world. We will start sending ping command periodically (once every 1 second) and record the delay and the jitter. Once every two minutes, we will add another mote to the system and continue pinging the same mote. We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system. This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
\paragraph{Experiment 2: }
The second experience is meant to measure the time it takes to transform the packet. In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system. The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
\section{Data Gathering}
\paragraph{}
After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
\subsection{Experiment 1}
\paragraph{}
In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes. The experiment as a whole took 12 minutes and 720 pings we done. Table \ref{table:exp1} shows the result of the experiment.

\begin{table}[htbp]
    \begin{tabular}{lll}
    \hline
    number of motes & Delay (ms) & Jitter (ms) \\\hline
    1               & 87.5         & 1.2         \\ 
    2               & 88.1         & 2.0         \\
    3               & 88.3        & 2.7         \\
    4               & 88.8         & 3.1         \\
    5               & 90.0         & 3.3         \\
    6               & 89.5         & 3.5         \\
    \end{tabular}
    \caption{Experiment 1 delay and jitter}
    \label{table:exp1}
\end{table}
\paragraph{}
For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row. The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval. The result is displayed in the table \ref{table:exp1}.

\subsection{Experiment 2}
\paragraph{}
The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system. The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient. Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.

@@

The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds. Depending on the machine's load.

@@

\subsection{Experiment 1}

@@

 nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}. To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}. 
 The good news in experiment 1 is that the privatization of packets brings benefit to the network. Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result. 
 \subsection{Experiment 2}
 The Packet transformation process is a component that shows excellent results. Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 2 but the results shows otherwise and that the component is not to be put responsible for such important delay seen in experiment 1.

@@

\backmatter

\addcontentsline{toc}{chapter}{Bibliography}
\renewcommand{\bibname}{References}
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}

\appendix
\chapter{How to install the system}
\section{Hardware needed}
In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services. The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi. A smartphone or a separate computer with Android emulator may de used.
